Homework 2

### 2016-10-18 Statistical Learning/Homework2/assignment.tex
Add an assignment

\problem{1}{Explain the relationship between spectral clustering, normalized spectral clustering and graph cut.}
\solution{Solution}{
We assume $k=2$ in the following explanation. \n
For spectral clustering, we already know that the eigenvector $u_2$ corresponding to the second smallest eigenvalue $\lambda_2$ of the Laplacian matrix of the graph is the solution for the minimization
$$ \min_{\tbf{f}} \ \tbf{f}^\top\tbf{L}\tbf{f},$$
$$s.t. \quad \tbf{f}^\top \tbf{f} = 1, \tbf{f}^\top\tbf{1} = 0$$
In fact, if we define 
$$G(f) = G(f_1,\cdots, f_n) = f\tr\tbf{L}f - \lambda_1 (f\tr f-1) - \lambda_2 f\tr\tbf{1} $$ $$ =  \frac{1}{2}\sum_{i,j=1}^{n}w_{ij}(f_i-f_j)^2-\lambda_1(\sum_{i=1}^{n}f_i^2-1)-\lambda_2\sum_{i=1}^n f_i$$
Then according to Lagrange Theorm, we let
$$\frac{\partial G}{\partial f_i} = \sum_{j=1}^n (f_i-f_j) - 2\lambda_1 f_i - \lambda_2 = 0, \quad  i=1:n$$
If we add the $n$ equations, we get
$$2\lambda_1\sum_{i, j=1}^n f_i + n\lambda_2=0$$
Since $f\tr\tbf{1} = 0$, then $\lambda_2 = 0$. So
$$\sum_{j=1}^n w_{ij}f_{j} = \lambda f_i, \quad i=1:n,$$
which means $f$ is an eigenvector of $\tbf{L}$.\n
For Min Cut, if we define $\tbf{f} = (f_1, f_2\cdots f_n)^\top$
}

% \problem{2}{Cluster $Data_i.csv$ with K-means, judge the number of clusters, and compare differences between different evaluating methods.}
% \solution{Solution}{
% The code of K-means, evaluating and test scripts can be found from attachments(kmeans.R, evaluate.R, hw1.2.R).
% }
% \solution{Result}{
% For Data1.csv, we cluster with k = 3. The result is shown as below.
% \begin{figure}[htbp]
% \centering
% \includegraphics[width = 15cm]{pics/hw2_data1_cluster.eps}
% \caption{The clusters of Data1.csv with k = 3}
% \label{clusters of data1}
% \end{figure}
% \\[1mm]

>>>>

\problem{1}{Explain the relationship between spectral clustering, normalized spectral clustering and graph cut.}
\solution{Solution}{
We assume $k=2$ in the following explanation. \n
For spectral clustering, we already know that the eigenvector $u_2$ corresponding to the second smallest eigenvalue $\lambda_2$ of the Laplacian matrix of the graph is the solution for the minimization
$$ \min_{\tbf{f}} \ \tbf{f}^\top\tbf{L}\tbf{f},$$
$$s.t. \quad \tbf{f}^\top \tbf{f} = 1, \tbf{f}^\top\tbf{1} = 0$$
In fact, if we define 
$$G(f) = G(f_1,\cdots, f_n) = f\tr\tbf{L}f - \lambda_1 (f\tr f-1) - \lambda_2 f\tr\tbf{1} $$ $$ =  \frac{1}{2}\sum_{i,j=1}^{n}w_{ij}(f_i-f_j)^2-\lambda_1(\sum_{i=1}^{n}f_i^2-1)-\lambda_2\sum_{i=1}^n f_i$$
Then according to Lagrange Theorm, we let
$$\frac{\partial G}{\partial f_i} = \sum_{j=1}^n (f_i-f_j) - 2\lambda_1 f_i - \lambda_2 = 0, \quad  i=1:n$$
If we add the $n$ equations, we get
$$2\lambda_1\sum_{i, j=1}^n f_i + n\lambda_2=0$$
Since $f\tr\tbf{1} = 0$, then $\lambda_2 = 0$. So
$$\sum_{j=1}^n w_{ij}f_{j} = \lambda f_i,\quad \lambda = \sum_{j=1}^{n}w_{ij}-2\lambda_1,\quad i=1:n, $$
which means $f$ is an eigenvector of $\tbf{L}$.\n
For Min Cut, if we define $\tbf{f} = (f_1, f_2\cdots f_n)^\top\in\mbb{R}^n$,
$$\bra{f_i}{1}{-1}{if\ v_i\in\mbb{A}}{if\ v_i\in\mbb{\bar{A}}}$$
Then
$$\bra{(f_i-f_j)^2}{4}{0}{i\in\mbb{A},j\in\mbb{\bar{A}}}{else}$$
So 
$$MCut(A_1,\cdots,A_k) = \sum_{i=1}^{k}cut(A_i, \bar{A_i}) = \frac{1}{4}f\tr\tbf{L}f$$
We can minimize this with the $f$ from spectral clustering. \n
For Ratio Cut, we can define $f$ as
$$\bra{f_i}{\sqrt{\vt{\bar{A}}/\vt{\bar{A}}}}{-\sqrt{\vt{A}/\vt{\bar{A}}}}{v_i\in\mbb{A}}{v_i\in\mbb{\bar{A}}}$$
Then
$$\bra{(f_i-f_j)^2}{\frac{\vt{A}}{\vt{\bar{A}}}+\frac{\vt{\bar{A}}}{\vt{A}}+2}{0}{v_i, v_j\ in \ the \ same \ part}{else}$$
So 
$$RCut(A, \bar{A}) = \frac{1}{2}\frac{\vt{A}\vt{\bar{A}}}{(\vt{A}+\vt{\bar{A}})^2}\frac{1}{\vt{A}+\vt{\bar{A}}}\sum_{i, j}(f_i-f_j)^2w_{ij}=\frac{f\tr\tbf{L}f}{2n^2}$$
For normalized spectral clustering, we have
$$L_{sym} = I - D^{-1/2}WD^{-1/2} = D^{-1/2}LD^{-1/2}$$
and normalized cut can be done using $g^*$, the Fiedler vector of $D^{-1/2}LD^{-1/2}.$ \\[1cm]
When $k \geq 3$, we can find $k$ points $x_i$ in $\mbb{R}^k$, s.t. $\vt{x_i-x_j} = \vt{x_k - x_l}$, for each $i, j, k, l$. \\
Then define $f_i = x_i$, and it is same with $k = 2$.
}

\problem{2}{Cluster Data.csv with spectral clustering, construct three different graphs and compare the difference with result of K-means.}
\solution{Solution}{
For Data1.csv, the result is shown as follows.
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_2_data1_kmeans.eps}
\caption{Kmeans}
\label{Kmeans for data 1}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Full-connect1]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_2_data1_full_1.eps}}\hfill
\subfloat[Full-connect2]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_2_data1_full_2.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Epsilon = 0.1]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_epsilon_01.eps}}\hfill
\subfloat[Epsilon = 1]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_epsilon_1.eps}}\hfill
\subfloat[Epsilon = 3]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_epsilon_3.eps}}	
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Directed, n = 3]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_directed_3.eps}}\hfill
\subfloat[Directed, n = 5]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_directed_5.eps}}\hfill
\subfloat[Directed, n = 10]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_directed_10.eps}}	
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Undirected, n = 3]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_undirected_3.eps}}\hfill
\subfloat[Undirected, n = 5]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_undirected_5.eps}}\hfill
\subfloat[Undirected, n = 20]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data1_undirected_20.eps}}	
\end{figure}
\\[1cm]
For Data3.csv, the result is shown as follows.
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_2_data3_kmeans.eps}
\caption{Kmeans}
\label{Kmeans for data 3}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Full-connect1]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_2_data3_full_1.eps}}\hfill
\subfloat[Full-connect2]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_2_data3_full_2.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[Epsilon = 0.1]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data3_epsilon_0_1.eps}}\hfill
\subfloat[Epsilon = 2]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data3_epsilon_2.eps}}\hfill
\subfloat[Epsilon = 5]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_2_data3_epsilon_5.eps}}	
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_2_data3_directed_3.eps}
\caption{directed, n = 3}
\end{figure}
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_2_data3_undirected_3.eps}
\caption{Undirected, n = 3}
\end{figure} \\
From the results we can know that for these groups, spectral clustering is not always better than kmeans. The parameters in spectral clustering are also important.
}
\newpage

\problem{3}{Cluster circles, curves with spectral clustering}
\solution{Solution}{
The results of circles.csv is as follows.
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_3_circle_full.eps}
\caption{full connection}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[epsilon = 1]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_epsilon_1.eps}}\hfill
\subfloat[epsilon = 2]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_epsilon_2.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[directed, n = 5]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_directed_5.eps}}\hfill
\subfloat[directed, n = 10]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_directed_10.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[undirected, n = 2]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_undirected_2.eps}}\hfill
\subfloat[undirected, n = 5]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_circle_undirected_5.eps}}
\end{figure}
\\
The result of curves.csv is shown as follows.
\begin{figure}[htbp]
\centering
\includegraphics[width = 10cm]{pics/hw2_3_curve_full.eps}
\caption{full connection}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[epsilon = 1]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_epsilon_1.eps}}\hfill
\subfloat[epsilon = 3]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_epsilon_3.eps}}\hfill
\subfloat[epsilon = 10]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_epsilon_10.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[directed, n = 10]{ %
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_directed_10.eps}}\hfill
\subfloat[directed, n = 20]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_directed_20.eps}}\hfill
\subfloat[directed, n = 30]{
	\includegraphics[width = 0.3\textwidth]{pics/hw2_3_curve_directed_30.eps}}
\end{figure}
\begin{figure}[htbp]
\centering
\subfloat[undirected, n = 5]{ %
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_curve_undirected_5.eps}}\hfill
\subfloat[undirected, n = 10]{
	\includegraphics[width = 0.45\textwidth]{pics/hw2_3_curve_undirected_10.eps}}
\end{figure}
}

% \begin{figure}[htbp]
% \centering
% \includegraphics[width = 10cm]{pics/hw2_data1_cluster.eps}
% \caption{The clusters of Data1.csv with k = 3}
% \label{clusters of data1}
% \end{figure}
% \\[1mm]
% \problem{2}{Cluster $Data_i.csv$ with K-means, judge the number of clusters, and compare differences between different evaluating methodss.}
% \solution{Solution}{
% The code of K-means, evaluating and test scripts can be found from attachments(kmeans.R, evaluate.R, hw1.2.R).
% }
% \solution{Result}{
% For Data1.csv, we cluster with k = 3. The result is shown as below.
% \begin{figure}[htbp]
% \centering
% \includegraphics[width = 10cm]{pics/hw2_data1_cluster.eps}
% \caption{The clusters of Data1.csv with k = 3}
% \label{clusters of data1}
% \end{figure}
% \\[1mm]
